---
title: "Public Pilot IRS Analysis"
author: "A Martin"
date: "2024-10-30"
output: pdf_document
---

This document provides the code to do the descriptive analysis in the first few paragraphs of the results section of the below manuscript:

Impact of late rainy season indoor residual spraying on holoendemic malaria transmission: a cohort study in northern Zambia

The code for the regresion models is in a separate stata.do file.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS Manuscript/")
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(plyr)
library(dplyr)
library(tidyr)
library(lubridate)
library(readxl)
library(table1)
library(mice)
library(ggplot2)
library(sjPlot)
library(geepack)
library(lme4) 
library(nlme)
library(glmmTMB)
library(gt)
library(kableExtra)
library(REDCapR)
library(RCurl)
library(geepack)
library(lme4) 
library(nlme)
library(splines)
library(optimx)
library(marginaleffects)
library(margins)
library(geomtextpath)
library(Hmisc) # load this last so that labels stick when subsetting
```

# Prep
Import datasets
```{r import datasets}
primary <- read.csv("/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS/Drafts/July 2024 JID Submission/Final code and datasets/primary.csv")
baseline_IRS <- read.csv("/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS/Drafts/July 2024 JID Submission/Final code and datasets/baseline_IRS.csv")
IRS_cov <- read.csv("/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS/Drafts/July 2024 JID Submission/Final code and datasets/IRS_cov.csv")
entomology <- read.csv("/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS/Drafts/July 2024 JID Submission/Final code and datasets/entomology.csv")
anopheline <- read.csv(file="/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS/Drafts/July 2024 JID Submission/Final code and datasets/anopheline.csv") 

#Filter baseline observations by taking the first observation for each individual
primary_base <- primary %>%
  filter(isparticipating==1) %>%
  group_by(partid) %>%
  slice(1) 
```


Run function to product p-values
```{r pvalue FUNCTION}
# Code source: https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html#example-a-column-of-p-values
# Updated the categorical test to use Fisher's exact since sample is small.
library(table1)
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- fisher.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
    
}
```

# Results paragraph 1
Household count
```{r hh count}
## Count of unique households
unique_hh <- primary %>%  # Build data frame with numbers of unique households who participated at least once
  filter(isparticipating==1) %>%
  group_by(hhnumber) %>%
  summarise(count = n_distinct(hhnumber))
nrow(unique_hh)

```
"Three hundred and eighty-two individuals from 64 households were enrolled."
Individual unique count comes from Table 1, below

Individuals per household
```{r indiv per hh}
summary(primary_base$TV.hhmemcount)

## Distribution of visits per person, by pre-post, by arm

#Overall
visit_tot <- primary %>% 
    filter(isparticipating==1) %>% # Removing missing
    group_by(partid) %>% 
    dplyr::summarise(visit_total = n())


summary(visit_tot$visit_total)
```
"Each household had a median of six occupants (IQR 4-8), who on average, were available and participated in 3.4 of six total possible visits."

Age distribution (years) and sex (proportion biologically male)
```{r age years and sex}
summary(primary_base$B.ageyears)

summary(primary_base$B.sex)[4] # proportion male
1- summary(primary_base$B.sex)[4] # proportion female
```
"The median age was 14 years (IQR 6-24) and the majority (58%) were female (Table 1). "

Net usage by age category
```{r net use by age}
table(primary$B.agecat.factor)
useagecounts<- table(primary$B.agecat.factor, primary$TV.netsleepln)
useageprop<- useagecounts[,2]/(useagecounts[,1]+useagecounts[,2])
useagecounts
useageprop
1-useageprop #nonusers

```
"Bed net usage differed by age group and was 92%, 62%, 50%, and 71% in those younger than one year, younger than five years, 6 – 16 years, and 17 years and older, respectively."

"Baseline vector counts and parasite prevalence by all measures were balanced across arms, although there were cluster-level (i.e. not sprayed inland, sprayed inland, not sprayed lakeside, sprayed lakeside) differences in RDT and microscopy prevalence (ANOVA test p-values = 0.012 and <0.01, Figure S1) and household vector counts (p-value < 0.01, Figure 2)."

```{r table 1 }
library(htmlTable)
library(tableone)

# Create dataset and add labels for table
dat <-primary_base %>% select(arm, B.ageyears , B.sex, B.residediffhome , B.mal.cause.know , TV.hlthlastmonth , TV.employmentyn , B.irscat, B.irsoctyn, TV.netsleepln , TV.nettimeown , TV.outnight, TV.hhmemcount , B.hheduclevel , hhhavebednet , hhbednetcnt , B.hheave , B.hhwall , B.hhrooftype, B.hhwatersrc , hhanimalspresent , B.distc , B.distr , B.dists1 , B.dists2 , B.dists3 , B.dists4 , B.dists5 , B.dists6, rdtresults, microresults, qpcrcall)
label(dat$B.ageyears)<- "Age (y)"
label(dat$B.sex)      <- "Sex"
label(dat$B.residediffhome)<- "Reside elsewhere for part of year"
label(dat$B.hhwatersrc)<- "Water Source"
label(dat$B.mal.cause.know)<- "Knows mosquito bites cause malaria"
label(dat$B.sex)      <- "Sex"
label(dat$B.hheduclevel)<- "HH Head Level of Education"
levels(dat$B.hheduclevel)=c("Primary","Secondary","Higher")
label(dat$hhhavebednet)= "House has bednet"
label(dat$B.irsoctyn)<- "Received IRS October 2022"
label(dat$TV.netsleepln)<- "Slept under bednet last night"
label(dat$TV.outnight)<- "Spends time away from home, outside after dusk"
label(dat$TV.hhmemcount)<- "Number of people in household"
label(dat$TV.hlthlastmonth) <- "Been to health facility in last month"
label(dat$TV.employmentyn) <- "Employed"
label(dat$TV.netsleepln) <- "Slept under bednet last night"
label(dat$TV.nettimeown) <- "Time owning net"
label(dat$TV.outnight) <- "Spends time outdoors after sundown"
label(dat$hhbednetcnt)<- "Total nets owned by household"
label(dat$B.hhwatersrc)<- "Household water source"
label(dat$hhanimalspresent)<- "Household owns animals"
#label(dat$hhappliance1)<- "Household owns radio"
#label(dat$hhappliance2)<- "Household owns TV"
#label(dat$hhappliance3)<- "Household owns refridgerator"
#label(dat$hhappliance4)<- "Household owns phone"
#label(dat$hhappliance5)<- "Household owns solar panels"
label(dat$B.distc)<- "Distance from health center (km)"
label(dat$B.distr)<- "Distance from road (km)"
label(dat$B.dists1)<- "Category 1 (km)"
label(dat$B.dists2)<- "Category 2 (km)"
label(dat$B.dists3)<- "Category 3 (km)"
label(dat$B.dists4)<- "Category 4 (km)"
label(dat$B.dists5)<- "Category 5 (km)"
label(dat$B.dists6)<- "Category 6 (km)"
label(dat$microresults)<- "RDT Positive"
label(dat$rdtresults)<- "Microscopy Positive"
label(dat$qpcrcall)<- "qPCR Positive"

# Modify factor levels for categorical variables
dat_cat <- dat
dat_cat$B.hhwall <- factor(dat_cat$B.hhwall,
                                  levels = c(1, 2, 3, 4),
                                  labels = c("Mud", "Concrete", "Wood", "Other"))
dat_cat$B.hhrooftype <- factor(dat_cat$B.hhrooftype,
                               levels = c(1, 2),
                               labels = c("Thatched", "Metal sheets"))
dat_cat$B.hhwatersrc <- factor(dat_cat$B.hhwatersrc,
                               levels = c(1, 2, 4),
                               labels = c("Piped water", "Well", "Other"))
dat_cat$B.hheduclevel <- factor(dat_cat$B.hheduclevel,
                               levels = c(1, 2),
                               labels = c("Primary", "Secondary"))
dat_cat$TV.nettimeown <- factor(dat_cat$TV.nettimeown,
                               levels = c(1, 2, 3),
                               labels = c("<1 month","1 month - 1 year","2 - 5 years"))
dat_cat$B.irscat <- factor(dat_cat$B.irscat,
                               levels = c(0, 1, 2, 3),
                               labels = c("No IRS","Oct. Only","April Only", "Both Oct. & April"))
label(dat_cat$B.hhwatersrc)<- "Water Source"
label(dat_cat$B.hhrooftype)<- "Roof type"


# Generate the table 1 using tableone with modified categorical variables
table1 <- CreateTableOne(
  vars = c("rdtresults", "microresults", "qpcrcall","B.ageyears", "B.sex", "B.residediffhome", "B.mal.cause.know", "TV.hlthlastmonth",
           "TV.employmentyn","B.irsoctyn", "hhhavebednet", "hhbednetcnt", "TV.nettimeown", "TV.netsleepln",
           "TV.outnight", "TV.hhmemcount", "B.hheduclevel",  "B.hheave",
           "B.hhwall", "B.hhrooftype", "B.hhwatersrc", "hhanimalspresent",
           "B.distc", "B.distr", "B.dists1", "B.dists2", "B.dists3", "B.dists4", "B.dists5",
           "B.dists6"),
  strata = "arm",
  data = dat_cat,
  addOverall = TRUE
)
t1<-print(table1, varLabels=TRUE)
```

Baseline vector and parasite measures
```{r baseline vector and parasite}
# Baseline differences in parasite prevalence
## RDT
rdt.chisq <- primary %>% dplyr::filter(studytime==0) %>% dplyr::filter(!is.na(B.Zone))
chisq.test(table(rdt.chisq$rdtresults, rdt.chisq$arm)) # difference by arm at baseline
chisq.test(table(rdt.chisq$rdtresults, rdt.chisq$B.Zone)) # difference by cluster at baselineff at baseline
## Micro
micro.chisq <- primary %>% dplyr::filter(studytime==0) %>% dplyr::filter(!is.na(B.Zone))
chisq.test(table(micro.chisq$microresults, micro.chisq$arm)) # difference by arm at baseline
chisq.test(table(micro.chisq$microresults, micro.chisq$B.Zone)) # difference by cluster at baseline
## qPCR
qpcr.chisq <- primary %>% dplyr::filter(studytime==0) %>% dplyr::filter(!is.na(B.Zone))
chisq.test(table(qpcr.chisq$qpcrcall, qpcr.chisq$arm)) # difference by arm at baseline
chisq.test(table(qpcr.chisq$qpcrcall, qpcr.chisq$B.Zone)) # difference by cluster at baseline

# Baseline differences in vector counts
anova.ento <- entomology %>% filter(studytime==0)
summary(aov(mcolanophelesf ~ B.Zone, data =anova.ento)) # test for difference at baseline across cluster
summary(aov(mcolanophelesf ~ arm, data =anova.ento)) # test for difference at baseline across arm


```
"Baseline vector counts and parasite prevalence by all measures were balanced across arms, although there were cluster-level (i.e. not sprayed inland, sprayed inland, not sprayed lakeside, sprayed lakeside) differences in RDT and microscopy prevalence (ANOVA test p-values = 0.012 and <0.01, Figure S1) and household vector counts (p-value < 0.01, Figure 2)."

# Spray coverage
October 2021
```{r spray october}
nrow(baseline_IRS) # Number of households surveyed in March 2022
hhbyarm<- table(baseline_IRS$arm)
hhbyarm[2] # total households in sprayed arm
hhbyarm[1] # total households in unsprayed arm
spraystatus <-table(baseline_IRS$irs_spray, baseline_IRS$arm)
spraystatus[2,2] # Number of households in sprayed arm sprayed
spraystatus[2,1] # Number of households in unsprayed arm sprayed

spraystatus[2,2] /hhbyarm[2] # Number of households in sprayed arm sprayed / total households in sprayed arm
spraystatus[2,1] /hhbyarm[1]# Number of households in unsprayed arm sprayed / total households in unsprayed arm
```
"Of the 50 households enrolled in March, 49 verified their October 2021 IRS campaign spray status. Only 34% (8/26) of intervention households and 36% (9/23) of control households reported having received IRS in October 2021 when the annual IRS campaign was conducted. "

April 2022
$100 * Sprayed/Sampled$
```{r spray coverage}
IRS_cov %>% 
    group_by(Zone) %>%
  summarise(Coverage = round(100*mean(irs_ntotal/5,na.rm = TRUE),2),
            Sampled= 5*n()) # 4 neighbors + sampled household = 5

```

# Epidemiological results
```{r rdt prev across}
## RDT Period prevelance
primary.part.pos <- primary %>% filter(rdtresults==1)
length(unique(primary.part.pos$partid)) / length(unique(primary.part.$partid)) # RDT period prevalence

# Microscopy incidence - use total failures and analysis time at risk from stata stset
314/33988 #daily
365*314/33988 # annual
1000 * 365*314/33988 # per 1000 person years

# qPCR incidence - use total failures and analysis time at risk from stata stset
489/32914 #daily
365*489/32914 # annual
1000 * 365*489/32914 # per 1000 person years
```
"From March to August 2022, the RDT period-prevalence was 71.7% and the incidence was 3372 and 5422 infections per 1000-person-years by microscopy and qPCR, respectively."

Average number of positive RDT among those who tested positive at least once
```{r positive RDTs}
nrow(primary.part.pos)/length(unique(primary.part.pos$partid)) # average number of positive tests per unique individual testing positive RDT

primary.part.micropos <- primary %>% filter(microresults==1 & lasexmicrltr>=0)
nrow(primary.part.micropos)/length(unique(primary.part.micropos$partid)) # average number of positive tests per unique individual testing positive MICRO

primary.part.pcropos <- primary %>% filter(qpcrcall==1)
nrow(primary.part.pcropos)/length(unique(primary.part.pcropos$partid)) # average number of positive tests per unique individual testing positive QPCR

```
"Individuals ever-positive (those who tested positive at least once) by RDT, microscopy, or qPCR, were positive by the corresponding diagnostic test more than once across the study period – on average, 2.4, 2.0, and 2.3 times respectively."

Regression analysis code in Stata do file.

# Entomological results
```{r ento}
tapply(entomology$mcolanophelesf, list(entomology$ym), sum) # monthly household vector counts

```
Human biting rate, infectivity rates, and EIR
```{r hbr and csp}
#Human biting rate calculations
hbr_avg <-entomology %>% 
  dplyr::select(ym, studytime, B.arm, TV.hbr) %>%
  summarise(Median = median(TV.hbr, na.rm=TRUE),
         Mean = mean(TV.hbr, na.rm=TRUE)
  )

hbr_arm <-entomology %>% 
  ungroup() %>%
  dplyr::select(ym, studytime, B.arm, TV.hbr) %>%
  dplyr::group_by(ym, studytime, B.arm) %>%
  dplyr::summarise(Median = median(TV.hbr, na.rm=TRUE),
         Mean = mean(TV.hbr, na.rm=TRUE),
  )

hbr_arm <- hbr_arm %>% mutate(studytime =ifelse(ym == "22-Mar",0,
                                       ifelse(ym == "22-Apr",1,
                                          ifelse(ym == "22-May",2,
                                             ifelse(ym == "22-Jun",3,
                                                    ifelse(ym == "22-Jul",4,
                                                           ifelse(ym == "22-Aug",5,"NA")))))))


hbr_zone <-entomology %>% 
  ungroup() %>%
  dplyr::select(ym, studytime, B.Zone, TV.hbr) %>%
  dplyr::group_by(ym, studytime, B.Zone) %>%
  dplyr::summarise(Median = median(TV.hbr, na.rm=TRUE),
         Mean = mean(TV.hbr, na.rm=TRUE)
  )

hbr_zone <- hbr_zone %>% mutate(studytime =ifelse(ym == "22-Mar",0,
                                       ifelse(ym == "22-Apr",1,
                                          ifelse(ym == "22-May",2,
                                             ifelse(ym == "22-Jun",3,
                                                    ifelse(ym == "22-Jul",4,

                                                                                                                      ifelse(ym == "22-Aug",5,"NA")))))))

# Combine datasets to get species specific hbr
spec_hh <- anopheline %>%
  group_by(hh_number, month, an_morph_id) %>%
  dplyr::summarise(n =n())
spec_hh_wide <- spread(spec_hh, an_morph_id, n) %>% mutate(month = tolower(month)) %>% dplyr::rename(hhnumber=hh_number)
entomology1 <- entomology %>%
  select(hhnumber, ym, TV.numsleep) %>% # merge with hh data
  mutate(month = ifelse(ym == "22-Mar", "march",
                        ifelse(ym == "22-Apr", "april",
                        ifelse(ym == "22-May", "may",
                        ifelse(ym == "22-Jun", "june",
                        ifelse(ym == "22-Jul", "july",
                               ifelse(ym == "22-Aug", "august", "NA"
                        )))))))
hbr_spec <- merge(entomology1, spec_hh_wide, by = c("hhnumber", "month"))
hbr_spec <- hbr_spec %>% mutate(hbr_gam = gambiae/TV.numsleep,
                                hbr_fun = funestus/TV.numsleep)
hbr_spec_month <- hbr_spec %>%
  dplyr::group_by(month) %>%
  summarise(Median_gam = median(hbr_gam, na.rm=TRUE),
            Mean_gam = mean(hbr_gam, na.rm=TRUE),
            Median_fun = median(hbr_fun, na.rm=TRUE),
         Mean_fun = mean(hbr_fun, na.rm=TRUE)
  )

hbr_spec_overall <- hbr_spec %>%
   summarise(Median_gam = median(hbr_gam, na.rm=TRUE),
            Mean_gam = mean(hbr_gam, na.rm=TRUE),
            Median_fun = median(hbr_fun, na.rm=TRUE),
         Mean_fun = mean(hbr_fun, na.rm=TRUE)
  )

# Circumsporozoite protein presence (infectivity) 
csp_arm<- anopheline %>%
  dplyr::rename(B.arm = arm) %>%
  dplyr:: group_by(studytime, B.arm) %>%
  dplyr::summarise(pos = sum(elisa_results, na.rm=TRUE),
            tot = n(),
            csp = pos/tot) 

csp_arm$csp <- ifelse(csp_arm$studytime==2 | csp_arm$studytime==4, NA, csp_arm$csp)

csp_zone <- anopheline %>%
  dplyr::rename(B.Zone = Zone) %>%
  dplyr:: group_by(studytime, B.Zone) %>%
  dplyr::summarise(pos = sum(elisa_results, na.rm=TRUE),
            tot = n(),
            csp = pos/tot) 

csp_zone$csp <- ifelse(csp_zone$studytime==2 | csp_zone$studytime==4, NA, csp_zone$csp)

csp_spec <-as.data.frame(anopheline %>%
  filter(elisa_results>=0) %>%
  dplyr:: group_by(an_morph_id) %>%
  dplyr::summarise(pos = sum(elisa_results, na.rm=TRUE),
            tot = n(),
            csp = pos/tot) )

csp_spec_mol <-as.data.frame(anopheline %>% # this filtering looks at molecular confirmation
  filter(elisa_results>=0) %>%
  dplyr:: group_by(final_id) %>%
  dplyr::summarise(pos = sum(elisa_results, na.rm=TRUE),
            tot = n(),
            csp = pos/tot) )

# Look at EIR estimates after combining above file with csp data
hbr_w_csp_arm <- merge(hbr_arm, csp_arm, by = c("studytime","B.arm"))
hbr_w_csp_EIR_arm <-  hbr_w_csp_arm %>% filter(!is.na(csp))
hbr_w_csp_EIR_arm$EIR_night = round(hbr_w_csp_EIR_arm$Mean*hbr_w_csp_EIR_arm$csp,2)
hbr_w_csp_EIR_arm$EIR_month = round(30 * hbr_w_csp_EIR_arm$Mean*hbr_w_csp_EIR_arm$csp,2)
hbr_w_csp_EIR_arm$EIR_year = round(365 * hbr_w_csp_EIR_arm$Mean*hbr_w_csp_EIR_arm$csp,2)

hbr_w_csp <- merge(hbr_zone, csp_zone, by = c("studytime","B.Zone"))
hbr_w_csp_EIR <-  hbr_w_csp %>% filter(!is.na(csp))
hbr_w_csp_EIR$EIR_night = round(hbr_w_csp_EIR$Mean*hbr_w_csp_EIR$csp,2)
hbr_w_csp_EIR$EIR_month = round(30 * hbr_w_csp_EIR$Mean*hbr_w_csp_EIR$csp,2)
hbr_w_csp_EIR$EIR_year = round(365 * hbr_w_csp_EIR$Mean*hbr_w_csp_EIR$csp,2)

hbr_w_csp_spec_EIR <- csp_spec %>% mutate(hbr = as.numeric(ifelse(an_morph_id=="funestus",hbr_spec_overall[4],
                                            ifelse(an_morph_id=="gambiae", hbr_spec_overall[2],"NA"))),
                                      EIR_night = round(hbr*csp,2),
                                      EIR_month = round(30 * hbr*csp,2),
                                      EIR_year = round(365 * hbr*csp,2)) %>%
  filter(an_morph_id == "funestus" |  an_morph_id == "gambiae")

hbr_w_csp_spec_EIR # EIR table
hbr_w_csp_spec_EIR[1,8] + hbr_w_csp_spec_EIR[2,8] # total EIR
```
"The EIR was highest in March,  and species-level annual EIR estimates from CDC light traps for An. funestus and An. gambiae, were 16.7 and 17.2, respectively for an overall EIR of 33.9 (Figure 3, Supplement, Table S1a)."

Regression analysis code in Stata do file.

#Figures 
```{r figure settings}
# Define color codes and settings for graphs
## Colors: https://www.nceas.ucsb.edu/sites/default/files/2020-04/colorPaletteCheatsheet.pdf
pallete2 <- c("slategray","darkolivegreen3")
pallete2.1 = c("gray", "black")
pallete4 <- c("darkgreen","orange","navy", "red")
pallete4age <- c("#4B0055", "#007094", "#00BE7D", "#FDE333")
pallete2feb <-c("navy", "red4")
fills =  c("darkblue","navy", "gray46", "black" )
lines3 = c(1,2,3,4)
pd <- position_dodge(1/10) # move them .5 to the left and right
```
Figure 1 is GIS Map

Figure 2
```{r figure 2 ento}
library(gridExtra)
lanoph_graph <- anopheline %>% select(an_morph_id, studytime, arm, Zone) %>%
  mutate(group = paste0(an_morph_id),
         order = paste0(Zone, studytime),
         ) %>%
  filter(an_morph_id=="funestus" |an_morph_id=="gambiae") %>% 
  mutate(group = ifelse(group=="funestus", "zfunestus", group)) %>%
  filter(!is.na(arm)) %>% #need to check this
  select(order, group)

#add empty rows for May and July
empty <- c("Lakeside, Sprayed6", "gambiae")
empty2 <- c("Lakeside, Not Sprayed6", "gambiae")
empty3 <- c("Inland, Sprayed6", "gambiae")
empty4 <- c("Inland, Not Sprayed6", "gambiae")

lanoph_graph1 <- as.data.frame(rbind(lanoph_graph, empty, empty2, empty3, empty4))
hbr_w_csp_EIR_arm <- as.data.frame(rbind(hbr_w_csp_EIR_arm, empty, empty2, empty3, empty4, empty5, empty6, empty7, empty8))

ento_plot1 <-ggplot(lanoph_graph1, aes(x=factor(order), fill=group)) +
  geom_bar(aes(),position=, color = "black") +
  labs(x='', y ='Total vector counts', fill="Species") +
  scale_fill_manual(values = c("orange","navy"), labels = c("An. gambiae","An. funestus"))+ 
  #ggtitle("Vector Species by Month") +
  scale_x_discrete(labels = c("Mar", "Apr", "May", "Jun", "Jul", "Aug"," ","Mar", "Apr", "May", "Jun", "Jul", "Aug"," ","Mar", "Apr", "May", "Jun", "Jul", "Aug"," ","Mar", "Apr", "May", "Jun", "Jul", "Aug"," "))+
  theme_classic() +
  theme(title = element_text(size=14, face="bold"),
    axis.text.x = element_text(angle = 90, size = 14),
    axis.text.y = element_text(size = 12),
  axis.title.x = element_text(size = 18),
  axis.title.y = element_text(size = 18),
  legend.title = element_text(size = 14),
  legend.text = element_text(size=12),
        axis.ticks.x = element_blank()) +
  geom_vline(aes(xintercept =7)) +
  geom_vline(aes(xintercept =14)) +
  geom_vline(aes(xintercept =21)) +
  geom_vline(aes(xintercept =28)) +
  annotate("text", x = 3.75, y = 900, label = "Not Sprayed, \nInland", size =5, fontface="bold", color = pallete4[1])+
  annotate("text", x = 10.5, y = 900, label = "Sprayed, \nInland", size =5, fontface="bold", color = pallete4[2])+
  annotate("text", x = 17.5, y = 900, label = "Not Sprayed, \nLakeside", size =5, fontface="bold", color = pallete4[3])+
  annotate("text", x = 24.5, y = 900, label = "Sprayed, \nLakeside", size =5, fontface="bold", color = pallete4[4])


ento_plot2 <- ggplot() + 
  stat_summary(data = lento_shorthh, aes(x = factor(studytime), y = mcolanophelesf, color = factor(B.Zone), group = factor(B.Zone)), geom = "line", fun = mean, size=1.5) +
  scale_color_manual(
    name = "Cluster",
    values = pallete4,
    labels = c("Not Sprayed, Inland", "Sprayed, Inland", "Not Sprayed, Lakeside", "Sprayed, Lakeside")) +
  #ylim(c(0, 30)) +
  scale_x_discrete(labels = c("March", "April","May","June","July", "August")) +
 labs(x = "", y = "Mean household nightly anoph. count") +
 theme_classic() +
  geom_textvline(label="Pilot Spray", xintercept=2,lwd=4,colour="blue", alpha = 0.35, hjust=.4, size = 6) +
  theme(title = element_text(size=14, face="bold"),
  axis.title.x = element_text(size = 18),
  axis.title.y = element_text(size = 18),
  legend.title = element_text(size = 14),
  legend.text = element_text(size=12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  legend.background = element_rect(linetype = 1, size = 0.25, colour = 1),
   legend.position="top")

#add empty rows for May and July
empty <- c(2,'Inland, Not Sprayed',NA, NA,NA,NA,NA,NA,NA,1.565,NA, NA)
empty2 <- c(2,'Lakeside, Not Sprayed',NA, NA,NA,NA,NA,NA,NA,.01,NA,NA)
empty3 <- c(2,'Lakeside, Sprayed',NA, NA,NA,NA,NA,NA,NA,.01,NA,NA)
empty4 <- c(2,'Inland, Sprayed',NA, NA,NA,NA,NA,NA,NA,1.585,NA,NA)
empty5 <- c(4,'Inland, Not Sprayed',NA, NA,NA,NA,NA,NA,NA,1.39,NA,NA)
empty6 <- c(4,'Lakeside, Not Sprayed',NA, NA,NA,NA,NA,NA,NA,0,NA,NA)
empty7 <- c(4,'Lakeside, Sprayed',NA, NA,NA,NA,NA,NA,NA,0,NA,NA)
empty8 <- c(4,'Inland, Sprayed',NA, NA,NA,NA,NA,NA,NA,1.19,NA,NA)

hbr_w_csp_EIR <- as.data.frame(rbind(hbr_w_csp_EIR, empty, empty2, empty3, empty4, empty5, empty6, empty7, empty8))
hbr_w_csp_EIR_arm <- as.data.frame(rbind(hbr_w_csp_EIR_arm, empty, empty2, empty3, empty4, empty5, empty6, empty7, empty8))

hbr_w_csp_EIR_arm.use <- hbr_w_csp_EIR_arm %>% filter(!is.na(B.arm)) %>% mutate(EIR_month = as.numeric(EIR_month))
 hbr_w_csp_EIR <- hbr_w_csp_EIR %>% filter(!is.na(B.Zone))
table1b_eir <- hbr_w_csp_EIR

EIR_month_zone <- ggplot(data = hbr_w_csp_EIR,
       aes(x = factor(studytime), y = as.numeric(EIR_month), group = factor(B.Zone), color = factor(B.Zone)), alpha=.5) +
    geom_point(size = 0.01) + geom_line(lwd=1.5)+
    ylim(c(0,6)) +
    labs(x='Month of 2022', y ='Estimated infectious bites per month') +
    #ggtitle("Estimated Entomological Innoculation Rates") +
    theme_classic() +
    scale_x_discrete(labels = c("March", "April","May","June", "July", "August")) +
  geom_textvline(label="Pilot Spray", xintercept=as.numeric(lento_shorthh$studytime[6]),lwd=4,colour="blue", alpha = 0.35, hjust=.8, size = 6) +
scale_colour_manual(name = 'Cluster', , values=pallete4, labels = c('Not Sprayed Inland','Sprayed Inland','Not Sprayed Lakeside','Sprayed Lakeside')) +
  theme(title = element_text(size=14, face="bold"),
  axis.title.x = element_text(size = 18),
  axis.title.y = element_text(size = 18),
  legend.title = element_text(size = 14),
  legend.text = element_text(size=12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14))+
  guides(color = FALSE)


  scale_colour_discrete(na.translate = F, name = "Cluster")  
    stat_summary(aes(group = factor(arm)), geom = "line", fun = median, width=0.1, size=1)
    #geom_errorbar(mapping=aes(x=ym, ymin=Mean-SD, ymax=Mean+SD), width=0.1, size=1, color="black")

entoplotall <-grid.arrange(
  top = textGrob("a", x = unit(0.05, "npc"), y = unit(0.95, "npc"), just = c("left", "top"), gp = gpar(fontsize = 20,fontface = "bold")),
  ento_plot2,
  right = textGrob("c", x = unit(0.05, "npc"), y = unit(0.99, "npc"), just = c("left", "top"), gp = gpar(fontsize = 20,fontface = "bold")),
  ento_plot1,
  bottom = textGrob("b", x = unit(0.05, "npc"), y = unit(0.95, "npc"), just = c("left", "top"), gp = gpar(fontsize = 20,fontface = "bold")),
EIR_month_zone, ncol =2
)


ggsave(file = "/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS/Drafts/July 2024 JID Submission/Final code and datasets/Public code outputs/entoplotall.png", width = 24, height = 16, entoplotall)
```

```{r Figure 3}
lmemhhpartpcr =  primary %>% filter(isparticipating=="1" & !is.na(qpcrcall)) # remove those who did not participate and have non na micro results
lmemhhpart =  primary %>% filter(isparticipating=="1" & !is.na(rdtresults)) # remove those who did not participate and have non na rdt results
lmemhhpartmicro =  primary %>% filter(isparticipating=="1" & !is.na(microresults)) # remove those who did not participate and have non na micro results



#https://journal.sjdm.org/14/141112a/summarySE.r # function
#troubleshooting: https://stackoverflow.com/questions/60403256/summaryse-function-throwing-error-cant-rename-columns-that-dont-exist-x-th
library(Rmisc)
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}

sumdat = summarySE(data = lmemhhpart, measurevar="rdtresults", groupvars=c("arm","ym"))
sumdat$ym <- factor(sumdat$ym, levels=c("22-Mar", "22-Apr", "22-May", "22-Jun", "22-Jul","22-Aug"))
sumdat$Outcome_Type = 1
sumdatmicro = summarySE(data = lmemhhpartmicro, measurevar="microresults", groupvars=c("arm","ym"))
sumdatmicro$ym <- factor(sumdatmicro$ym, levels=c("22-Mar", "22-Apr", "22-May", "22-Jun", "22-Jul","22-Aug"))
sumdatmicro$Outcome_Type = 2
sumdatqpcr = summarySE(data = lmemhhpartpcr, measurevar="qpcrcall", groupvars=c("arm","ym"))
pd <- position_dodge(1/10) # move them .5 to the left and right
sumdatqpcr$ym <- factor(sumdatqpcr$ym, levels=c("22-Mar", "22-Apr", "22-May", "22-Jun", "22-Jul","22-Aug"))
 sumdatqpcr$Outcome_Type = 3
# Make a master dataset for by month measurments
sumdat <- sumdat %>% dplyr::rename(result = rdtresults) #%>% filter(!is.na())
sumdatmicro <- sumdatmicro %>% dplyr::rename(result = microresults)
sumdatqpcr <- sumdatqpcr %>% dplyr::rename(result = qpcrcall)
sumdat_mas <- rbind(sumdat, sumdatmicro, sumdatqpcr)

# Master plot over time
shapes2 <- c(21, 22, 23, 25)
lines <- c(1,1,2,2,4,4,3,3)
breaks = 3 *c(0.2, 0.4, 0.6, 0.8, 1.0) # manually set these breaks based on values of log density
pallete2.1 = c("gray", "black")

convert_ratio2 <- function(x) {
  x /(1/3)
}

sumdat_mas$group = paste0(sumdat_mas$Outcome_Type,sumdat_mas$arm)
library(geomtextpath)
sumdat_mas <- sumdat_mas %>% filter(ym!="22-Sep")
figure1_master<- ggplot(sumdat_mas, aes(x=ym, y=result, color = as.factor(arm), fill = as.factor(arm),  group=group, linetype =as.factor(Outcome_Type), shape = as.factor(Outcome_Type))) +
  theme_classic() +
  geom_line() +
  geom_point(position=pd, size = 3) +
 	#geom_errorbar(aes(ymin=result-ci, ymax=result+ci), width=.1, position=pd) +
  #ggtitle("Epidemiological Outcomes") +
  xlab("Month of 2022") + ylab("Prevalence") +
  ylim(0,0.7)+
  geom_textvline(label="Pilot Spray", xintercept=2,lwd=4,colour="blue", alpha = 0.35, hjust=.1) + 
  scale_x_discrete(labels = c("March", "April", "May", "June", "July", "August")) +
  scale_fill_manual(values=pallete2.1, labels = c("Not Sprayed", "Sprayed"), name = "Arm") +
  scale_colour_manual(values=pallete2.1, labels = c("Not Sprayed", "Sprayed"), name = "Arm") +
  scale_linetype_manual(values = c(1,2,4), labels = c("RDT", "Microscopy", "qPCR"), name = "Outcome Type") +
  scale_shape_manual(values = c(16, 17, 18), labels = c("RDT", "Microscopy", "qPCR"), name = "Outcome Type") +  # Define shapes for each Outcome Type
  theme(title = element_text(size=14, face="bold"),
  axis.title.x = element_text(size = 12),
  axis.title.y = element_text(size = 12),
  legend.title = element_text(size = 12)) #+
  #scale_y_continuous(sec.axis = sec_axis(~convert_ratio2(.), name = "Log Parasites/Microliter", breaks = breaks, labels = breaks)) +
   #                  theme(axis.line.y.right = element_line(linetype=3),
    #                       legend.title = element_text(size=10))

ggsave(file = "/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS/Drafts/July 2024 JID Submission/Final code and datasets/Public code outputs/figure3.png", width = 10, height = 6, figure1_master)

```
Figure 4 in stata

Figure 5
```{r figure 5}
# Create the vertical forest plot for stata models
shapes5 <- c(23, 21, 24, 25, 22)
colors5 <- c("gray","black","black", "black" ,"black")

#Prepare the dataframe
rdt_vec_stata <- c(2, 0.98, 0.14, 6.95)
micro_vec_stata <- c(4, 0.31, 0.05,1.93)
qpcr_vec_stata <- c(3, 1.22,	0.76,1.95)
haz_vec_stata <- c(5, 0.477, 0.292 ,0.780)
ento_vec_stata <- c(1, 2.17, 0.87,5.43)

data_fp_plot_stata <-as.data.frame(rbind(ento_vec_stata, rdt_vec_stata, qpcr_vec_stata ,micro_vec_stata, haz_vec_stata )) 
colnames(data_fp_plot_stata) <- c("Outcome_Type", "Estimate", "Lower_CI", "Upper_CI")
data_fp_plot_stata$Time <- "m012345"
data_fp_plot_stata$Time <- factor(data_fp_plot_stata$Time, levels = c("m012345"), labels = c("Through August"))

fplotstata <- ggplot(data_fp_plot_stata, aes(x = Time, y = Estimate)) +
  geom_pointrange(aes(ymin = Lower_CI, ymax = Upper_CI, shape = factor(Outcome_Type), color = factor(Outcome_Type), fill= factor(Outcome_Type)), position = position_dodge2(w = 0.5), size = 1) +
  scale_shape_manual(values = shapes5, labels = c("Anopheles Counts", "RDT Prevalence", "qPCR Incidence","Microscopy Incidence", "Microscopy Hazard")) +
  scale_color_manual(values = colors5, labels = c("Anopheles Counts", "RDT Prevalence", "qPCR Incidence","Microscopy Incidence", "Microscopy Hazard")) +
  scale_fill_manual(values = colors5, labels = c("Anopheles Counts", "RDT Prevalence", "qPCR Incidence","Microscopy Incidence", "Microscopy Hazard")) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
  labs(y = "Change in sprayed / Change in unsprayed", shape = "Outcome", color = "Outcome", fill = "Outcome") +
  #scale_x_discrete(labels = c("May", "June", "July", "August")) +
  coord_cartesian(ylim = c(0,2.5)) +
  theme_classic() + # Gives white background
  theme(axis.text.y = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank())

fplotstata

ggsave("/Users/annemartin/Documents/PhD/ICEMR/Nchelenge/Pilot IRS/Drafts/July 2024 JID Submission/Final code and datasets/Public code outputs/figure5.jpg",width = 10, height = 6,fplotstata)
```
